The Radio Sky at Meter Wavelengths: $m$-Mode Analysis Imaging with the Owens Valley Long Wavelength Array

### 2017-07-26 paper.tex
progress towards first draft

@@

@@

\subsection{Gain Calibration}

The antenna complex gains are measured from a 7.2 hour track of data from when Cyg A and Cas are at
high elevations. The gains measured in this way are then used to calibrate the entire 28 hour
dataset. Cyg A and Cas A are -- by an order of magnitude -- the brightest point-like radio sources
in the northern hemisphere. Therefore the optimal time to solve for the interferometer's gain
calibration is when these sources are at high elevations. \todo{cite some statistics on how good
this calibration is? what do I quantify?}

\todo{at the moment we calibrated with the Baars spectrum for Cas. we should actually iterate once
and use the new spectrum that comes out of the beam fitting routine}

The flux scale is tied to the work of \citet{1977A&A....61...99B, 2012MNRAS.423L..30S,
2017ApJS..230....7P}.

\todo{tuesday}

\begin{figure*}[ht]
    \plottwo{figures/beam/source-tracks}{figures/beam/stokes-I-beam}
    \caption{
        Example Stokes-I beam model (from spw04).
    }
    \label{fig:beam}
\end{figure*}

\subsubsection{Cassiopeia A and Cygnus A}

Without removing bright sources from the data, sidelobes from bright sources will dominate the
variance in the image.  At 74 MHz Cyg A is a 15,000 Jy source \citep{2017ApJS..230....7P}. A
conservative estimate for the confusion limit at 74 MHz with a 10 arcminute beam is 500 mJy
\citep{2012RaSc...47.0K04L}. Therefore we require that Cyg A's sidelobes be at most of $-$45 dB down
from its peak flux to prevent Cyg A's sidelobes from dominating the variance in the image.
\todo{Measure Cyg A's sidelobe levels if it is not subtracted}

\todo{Show a before and after with the updated source models}

\todo{Give a plot with the updated source models}

The Sun can be trivially removed from any map of the sky by constructing the map using only data
taken at night. A map of the entire sky can be obtained by using observations spaced 6 months apart.
However the dataset used in this paper consists of 28 consecutive hours. Additionally while most
astronomical sources at these frequencies have negative spectral indices, the Sun has a large
positive spectral index. Therefore more care will need to be taken in subtracting the Sun at higher
frequencies than at lower frequencies.

\todo{describe sun removal -- this isn't entirely settled yet}

@@

\todo{compute B*v for panels (c) and (d) to show that the pick-up is pretty bad. my expectation is
that it is actually worse than the RFI, but I don't actually know for sure}

@@

\todo{discuss the possibility of declination dependent flux-scale errors in these maps}

The second case is more subtle. Sources are observed at a wide range of locations in the primary
beam of the antenna. The imaging process must reconcile all of these observations together. The beam
model essentially provides the instructions on how to do this. For example if at a time $t_1$ the
antenna gain towards a source is $g$ and at a later time $t_2$ the antenna gain towards that same
source (which has now moved) is $g/2$, then in order to correctly estimate the flux of the source
the observations from time $t_1$ need to be multiplied by $1/g$ but the observations from time $t_2$
need to be multiplied by $2/g$. A mistake in the beam model here will lead to a mistake in the
estimated flux.  However in general the amplitude and phase of the antenna response needs to be
known and so beam model errors can lead to observations being combined with incorrect phase as well.
This will lead to a degraded PSF in the final map. \todo{quick simulation of this?}

@@

This will tend to scatter power between $m$-modes. \todo{cite richard for this}

@@

\floattable
\begin{deluxetable}{ccccc}
    \tablecaption{Summary of the produced maps\label{tab:summary}}
    \tablehead{
        \colhead{\#} & \colhead{$\nu$ / MHz} & \colhead{$\Delta\nu$ / MHz} &
        \colhead{$\theta$ / arcmin} & \colhead{noise / Jy}
    }
    \startdata
        1 & 36.528 & 0.024 & & \\
        2 & 41.760 & 0.024 & & \\
        3 & 46.992 & 0.024 & & \\
        4 & 52.224 & 0.024 & & \\
        5 & 57.456 & 0.024 & & \\
        6 & 62.688 & 0.024 & & \\
        7 & 67.920 & 0.024 & & \\
        8 & 73.152 & 0.024 & & \\
    \enddata
\end{deluxetable}

>>>>

\author{Lincoln Greenhill}
\affiliation{Harvard}

\author{Hugh Garsden}
\affiliation{Harvard}

\author{Stephen Bourke}
\affiliation{Chalmers}

@@

\begin{deluxetable*}{ccccc}
    \tablecaption{Summary of the generated all-sky maps\label{tab:summary}}
    \tablewidth{0pt}
    \tablehead{
        \colhead{\#} & \colhead{$\nu$ / MHz} & \colhead{$\Delta\nu$ / MHz} &
        \colhead{$\theta$ / arcmin} & \colhead{noise / Jy}
    }
    \startdata
        1 & 36.528 & 0.024 & & \\
        2 & 41.760 & 0.024 & & \\
        3 & 46.992 & 0.024 & & \\
        4 & 52.224 & 0.024 & & \\
        5 & 57.456 & 0.024 & & \\
        6 & 62.688 & 0.024 & & \\
        7 & 67.920 & 0.024 & & \\
        8 & 73.152 & 0.024 & & \\
    \enddata
\end{deluxetable*}

The antenna complex gains are measured from a 22 minute track of data from when Cyg A and Cas are at
high elevations. The gains measured in this way are then used to calibrate the entire 28 hour
dataset. Cyg A and Cas A are -- by an order of magnitude -- the brightest point-like radio sources
in the northern hemisphere. Therefore the optimal time to solve for the interferometer's gain
calibration is when these sources are at high elevations. The calibration sky model consists only of
Cyg A and Cas A. The flux of Cyg A is anchored to the results of \citet{1977A&A....61...99B} while
the flux of Cas A is measured from the data itself (using a preliminary calibration).

\begin{figure*}[ht]
    \plottwo{figures/beam/source-tracks}{figures/beam/stokes-I-beam}
    \caption{
        (left panel) The tracks traced through the sky by the 8 brightest point sources visible to the
        OVRO-LWA.
        (right panel) The amplitude of the OVRO-LWA beam at 36.528 MHz as a result of the beam model
        fit using Zernike polynomials and the symmetry assumptions outlined in the text.
    }
    \label{fig:beam}
\end{figure*}

\subsubsection{Cygnus A and Cassiopeia A}

%\begin{deluxetable*}{ccccDCCD}
%    \tablecaption{Resolved source models for Cyg A and Cas A\label{tab:resolved-source-models}}
%    \tablewidth{0pt}
%    \tablehead{
%        \colhead{} & \colhead{} &
%            \colhead{} & \colhead{} &
%            \multicolumn2c{Amplitude} &
%            \colhead{Major FWHM} & \colhead{Minor FWHM} &
%            \multicolumn2c{Position angle} \\
%        \colhead{Source} & \colhead{\#} &
%            \colhead{RA} & \colhead{Dec} &
%            \multicolumn2c{(relative)} &
%            \colhead{(arcsec)} & \colhead{(arcsec)} &
%            \multicolumn2c{(deg)}
%    }
%    \decimals
%    \startdata
%        Cyg A & 1 & & & 1.0 & 128. & 22. & -74.503 \\
%              & 2 & & & 0.1477 & 183. & 141. & 43.449 \\[0.2cm]
%        Cas A & 1 & & & 1.0 & 0. & 0. & 0. \\
%              & 2 & & & 1.0 & 0. & 0. & 0. \\
%              & 3 & & & 1.0 & 0. & 0. & 0. \\
%    \enddata
%\end{deluxetable*}

Without removing bright sources from the data, sidelobes from bright sources will dominate the
variance in the image.  At 74 MHz Cyg A is a 15,000 Jy source \citep{2017ApJS..230....7P}. A
conservative estimate for the confusion limit at 74 MHz with a 10 arcminute beam is 500 mJy
\citep{2012RaSc...47.0K04L}. Therefore we require that Cyg A's sidelobes be at most of $-$45 dB down
from its peak flux to prevent Cyg A's sidelobes from dominating the variance in the image.

Armed with resolved source models for both Cyg A and Cas A, these two sources can be peeled from the
dataset to the required dynamic range.

%\todo{slice through Cyg A and Cas A to illustrate the removal process}
%\todo{Give a plot with the updated source models}

\subsubsection{Other Bright Sources}

Other bright sources including Vir A, Tau A, Her A, Hya A, Per B, and 3C 353 are also removed from
the visibilities prior to imaging. Because these sources are much fainter than Cyg A and Cas A, we
do not need resolved source models to be able to remove these sources from the visibilities without
residual sidelobes contaminating the image.

However, the ionosphere will cause these sources to scintillate and refract. So first the position
and flux of each source is measured separately in each channel and integration. Then the sources are
subtracted from the visibilities using the updated position and flux of the source. The brightest of
these sources (Vir A and Tau A) will be peeled using a direction dependent calibration when they are
at high elevations.

The sun can be trivially removed from any map of the sky by constructing the map using only data
taken at night. A map of the entire sky can be obtained by using observations spaced 6 months apart.
However the dataset used in this paper consists of 28 consecutive hours. Therefore we attempt to
suppress the contribution of the sun to the maps with the goal of suppressing its sidelobes.  The
sun is well-resolved by the OVRO-LWA and hence a somewhat detailed source model is needed. In fact
the optical depth $\tau=1$ surface of the sun changes with frequency and as a consequence a new
model is needed for each of the 8 channel maps produced in this paper. So while we could fit a
limited number of Gaussian components to Cyg A and Cas A, this is insufficient for the Sun.
Additionally while most astronomical sources at these frequencies have negative spectral indices,
the Sun has a large positive spectral index. Therefore more care will need to be taken in
subtracting the Sun at higher frequencies than at lower frequencies.

The strategy used for removing the Sun below 55 MHz involves fitting a shapelet
\citep{2003MNRAS.338...35R} model to the Sun and subtracting without the use of direction dependent
gains. The shapelet fitting is performed in the visibility space. Above 55 MHz a model is fit to the
Sun by minimizing the residuals after peeling (in the same way that models are obtained for Cyg A
and Cas A). The Sun is then peeled from each integration using direction dependent gains.

\subsection{Flux Scale}

\begin{figure*}[p]
    \gridline{
        \fig{figures/flux-scale/flux-cyga}{0.48\textwidth}{}
        \fig{figures/flux-scale/flux-vira}{0.48\textwidth}{}
    }
    \gridline{
        \fig{figures/flux-scale/flux-hyaa}{0.48\textwidth}{}
        \fig{figures/flux-scale/flux-perb}{0.48\textwidth}{}
    }
    \gridline{
        \fig{figures/flux-scale/flux-lyna}{0.48\textwidth}{}
        \fig{figures/flux-scale/flux-3c48}{0.48\textwidth}{}
    }
\end{figure*}
\begin{figure*}[p]
    \gridline{
        \fig{figures/flux-scale/flux-3c147}{0.48\textwidth}{}
        \fig{figures/flux-scale/flux-3c286}{0.48\textwidth}{}
    }
    \gridline{
        \fig{figures/flux-scale/flux-3c295}{0.48\textwidth}{}
        \fig{figures/flux-scale/flux-3c353}{0.48\textwidth}{}
    }
    \gridline{
        \fig{figures/flux-scale/flux-3c380}{0.48\textwidth}{}
    }
    \caption{
        This figure shows the measured fluxes (black points) of 11 sources plotted against the
        published spectra from \citet{2017ApJS..230....7P} (solid line), \citet{2012MNRAS.423L..30S}
        (dashed line), and \citet{1977A&A....61...99B} (dot-dash line).
    }
    \label{fig:flux-scale}
\end{figure*}

The flux scale of the data was anchored to the \citet{1977A&A....61...99B} spectrum of Cyg A during
gain calibration. However, gain calibration is also a function of the beam model and the spectrum
used for Cas A. Recent work by \citet{2012MNRAS.423L..30S} using archival date from the literature
and \citet{2017ApJS..230....7P} using the VLA has expanded the number of low-frequency radio sources
with calibrated flux measurements from one (just Cyg A) to eleven in total. However while the
\citet{2012MNRAS.423L..30S} flux scale is valid between 30 MHz and 300 MHz, the
\citet{2017ApJS..230....7P} flux scale is somewhat more limited because the lowest frequency
observations come from the VLA 4-band system. As a consequence the \citet{2017ApJS..230....7P} flux
scale is not valid below 50 MHz.

Figure~\ref{fig:flux-scale} shows a comparison between flux measurements made using the all-sky maps
from this work, and spectra from the aforementioned flux scales. Generally the OVRO-LWA flux
measurements agree to better than 10\% with the \citet{2012MNRAS.423L..30S} spectra. Below 50 MHz
there can be substantial departures with respect to the \citet{2017ApJS..230....7P} spectra (for eg.
3C 286, 3C 295, and 3C 380), but it is usually the case that we have much better agreement with the
\citet{2015A&A...582A.123H} spectra. This indicates that the \citet{2017ApJS..230....7P} spectra can
diverge from the truth rapidly below 50 MHz.

@@

@@

%\todo{compute B*v for panels (c) and (d) to show that the pick-up is pretty bad. my expectation is
%that it is actually worse than the RFI, but I don't actually know for sure}

@@

%\todo{discuss the possibility of declination dependent flux-scale errors in these maps}

The second case is more subtle. Sources are observed at a wide range of locations in the primary
beam of the antenna. The imaging process must reconcile all of these observations together. The beam
model essentially provides the instructions on how to do this. For example if at a time $t_1$ the
antenna gain towards a source is $g$ and at a later time $t_2$ the antenna gain towards that same
source (which has now moved) is $g/2$, then in order to correctly estimate the flux of the source
the observations from time $t_1$ need to be multiplied by $1/g$ but the observations from time $t_2$
need to be multiplied by $2/g$. A mistake in the beam model here will lead to a mistake in the
estimated flux.  However in general the amplitude and phase of the antenna response needs to be
known and so beam model errors can lead to observations being combined with incorrect phase as well.
This will lead to a degraded PSF in the final map.

%\todo{quick simulation of this?}

@@

This will tend to scatter power between $m$-modes \citep{richard_ionosphere_thoughts}.

@@

\begin{figure*}[p]
    \gridline{
        \fig{figures/channel-maps/spw04-glamour-shot-map-wiener-filtered-rainy-2048-galactic.pdf}
            {0.48\textwidth}{}
        \fig{figures/channel-maps/spw06-glamour-shot-map-wiener-filtered-rainy-2048-galactic.pdf}
            {0.48\textwidth}{}
    }
    \gridline{
        \fig{figures/channel-maps/spw08-glamour-shot-map-wiener-filtered-rainy-2048-galactic.pdf}
            {0.48\textwidth}{}
        \fig{figures/channel-maps/spw10-glamour-shot-map-wiener-filtered-rainy-2048-galactic.pdf}
            {0.48\textwidth}{}
    }
    \gridline{
        \fig{figures/channel-maps/spw12-glamour-shot-map-wiener-filtered-rainy-2048-galactic.pdf}
            {0.48\textwidth}{}
        \fig{figures/channel-maps/spw14-glamour-shot-map-wiener-filtered-rainy-2048-galactic.pdf}
            {0.48\textwidth}{}
    }
    \gridline{
        \fig{figures/channel-maps/spw16-glamour-shot-map-wiener-filtered-rainy-2048-galactic.pdf}
            {0.48\textwidth}{}
        \fig{figures/channel-maps/spw18-glamour-shot-map-wiener-filtered-rainy-2048-galactic.pdf}
            {0.48\textwidth}{}
    }
    \caption{
        Beautiful chanel maps.
    }
    \label{fig:channel-maps}
\end{figure*}
